Neural and Symbolic Spatiotemporal Optimization of Cosmological Particle-Mesh Simulation

### 2023-03-21 docs/papers/sto/sto.tex
FIXUP draft: edit dimensionless NN input factors
FIXUP draft: edit TODOs

\section*{TODO}
\begin{itemize}
\item Moving some of the following ongoing and completed tasks to the
    main text and add preliminary figures.
\item Implement the rest of the nonlinear scales, including the rms
    displacement with np.trapz of $P$ over $\ln k$, the $\ln a$ time
    derivatives, etc.\newline
    \textbf{Update}: code has been added.
\item Compare one pair of Gadget4 and \pmwd\ (with say 100 to 1000 time
    steps) results and make sure that they agree perfectly on the large
    scales. For this purpose the box size should be large enough (in
    linear regime), say 1 Gpc/$h$.\newline
    \textbf{Update}: for Sobol=6, boxsize=983Mpc, Gadget4 and \pmwd\ are visually
    consistent at the field level. To further check power spectrum. For odd mesh
    shapes (e.g. 1, 3, 5), pmwd and Gadget are perfectly consistent on large
    scales. \textbf{While for even mesh shapes (e.g. 2, 4), the powers look
    slightly different. To figure out this weird issue.}
\item Be careful with the non-deterministic ICs generated on GPUs.\newline
      \textbf{Update}: we have shown that ICs generated on GPUs are
      deterministic (but different from those on CPUs), generating ICs for both
      training data and training on GPUs might be the simplest solution.
\item Incorporate 3lpt branch; testing; find a\_start for different mass
  resolutions
\item For TO, probably necessary to first incorporate Oliver Hahn's
  recent improvement on FastPM integrator arXiv:2301:09655.
\item Compare various field and halo summary stats to compare sto with
  raw pmwd, deconvolved pmwd, and gadget test cases.
\item Non-integer multiple mesh shapes: for uniform particle grid, test
  (with double precision on CPUs) whether they start moving by artifact
  forces on cell scales, in 3 cases: no deconvolution, with CIC
  deconvolution, with SO.
\end{itemize}

@@

Relevant scale-independent dimensionless factors are
\begin{itemize}
\item $D(a)$: linear growth factor;
\item $f(a)$: linear growth rate $f \triangleq \d\ln D / \d\ln a$;
\item $\Omegam(a)$: matter density parameter at $a$;
\item $\d\ln H / \d\ln a$: time derivative of Hubble expansion;
\item $a$: scale factor;
\item $\Delta a$: time step size in $a$.
\end{itemize}

>>>>

\section*{TODO}
\begin{itemize}
\item Moving some of the following ongoing and completed tasks to the
    main text and add preliminary figures.
\item Implement the rest of the nonlinear scales, including the rms
    displacement with np.trapz of $P$ over $\ln k$, the $\ln a$ time
    derivatives, etc.\newline
    \textbf{Update}: code has been added.
\item Compare one pair of Gadget4 and \pmwd\ (with say 100 to 1000 time
    steps) results and make sure that they agree perfectly on the large
    scales. For this purpose the box size should be large enough (in
    linear regime), say 1 Gpc/$h$.\newline
    \textbf{Update}: for Sobol=6, boxsize=983Mpc, Gadget4 and \pmwd\ are visually
    consistent at the field level. To further check power spectrum. For odd mesh
    shapes (e.g. 1, 3, 5), pmwd and Gadget are perfectly consistent on large
    scales. \textbf{While for even mesh shapes (e.g. 2, 4), the powers look
    slightly different. To figure out this weird issue.}
\item Be careful with the non-deterministic ICs generated on GPUs.\newline
      \textbf{Update}: we have shown that ICs generated on GPUs are
      deterministic (but different from those on CPUs), generating ICs for both
      training data and training on GPUs might be the simplest solution.
\item Incorporate 3lpt branch; testing; find a\_start for different mass
  resolutions
\item Probably necessary to first incorporate List \& Hahn's recent
  improvement on FastPM integrator arXiv:2301:09655.
\item Compare sto DM and halo fields and summary stats with those of raw
  pmwd, deconvolved pmwd, and gadget, on the last 9 test cases.
\item Demonstrate capability of generalization, on some cases with
  extrapolating cosmological parameters, etc.
\item Non-integer multiple mesh shapes: for uniform particle grid, test
  (with double precision on CPUs) whether they start moving by artifact
  forces on cell scales, in 3 cases: no deconvolution, with CIC
  deconvolution, with SO.
\end{itemize}

@@

Relevant scale-independent dimensionless factors are
\begin{itemize}
\item $G_m(a) \triangleq D_m(a) / a^m$: linear growth given in
  suppression factors for $m \in \{1, 2\}$;
\item $\d\ln G_m / \d\ln a$: with $m=1$ case related to the growth rate
  $f \triangleq \d\ln D_1 / \d\ln a$;
\item $\Omegam(a)$: matter density parameter at $a$;
\item $\d\ln H / \d\ln a$: time derivative of Hubble expansion;
\item $\Delta\ln a$: time step size in $\ln a$.
\end{itemize}
